---
title: "Lab 1: Wrangling review, new functions, AND A MAP!"
author: "Marie Bouffard"
date: "1/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Attach packages

library(tidyverse)
library(here)
library(sf)
library(tmap)

```

Read in data
```{r}
#---------------------------------------
# read in sf tree data
#---------------------------------------

sf_trees <- read_csv(here("data","sf_trees","sf_trees.csv"))

#---------------------------------------
# Exploratory functions
#---------------------------------------

# view(sf_trees)
# summary(sf_trees)
# names(sf_trees)

```

### Part 1: Wrangling & ggplot review
```{r}

#---------------------------------------
# Find counts by legal status and wrangle
#---------------------------------------

# Using group_by %>% summarize %>% n
sf_trees %>% 
  group_by(legal_status) %>% 
  summarize(tree_count = n())

# Same thing + new dplyr functions
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(tree_count, n = 5) %>% 
  arrange(-tree_count)

```

Graph top 5
```{r}

ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col() +
  labs(y = "Tree count", x = "Legal Status") +
  coord_flip() +
  theme_minimal() 

```

Filter for permitted and MTA
```{r}

permitted_mta <- sf_trees %>% 
  filter(legal_status == "Permitted Site", caretaker == "MTA")

```

Filter for Blackwood Acacia trees trees and select for columns + Meet stringr
```{r}

# str_detect() is part of stringr package for working with strings and is useful for finding a specific string in a column

blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude)

# Graph locations
ggplot(data = blackwood_acacia, aes(x = longitude, y = latitude)) + 
  geom_point()

```

Use tidyr::separate() to separate the species column into two separate columns
```{r}

sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = " :: ")

```

Using tidyr::unite() to combine two columns
```{r}

id_status_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_COOL_")

```

### Part 2: MAPS!

Converting to spatial coordinates using st_as_sf()
```{r}

# Converting lat/lon to spatial points
blackwood_acacia_sp <- blackwood_acacia %>% 
  drop_na(longitude, latitude) %>% 
  st_as_sf(coords = c("longitude","latitude"))

# Set CRS
st_crs(blackwood_acacia_sp) = 4326

# Plot in ggplot using geom_sf
ggplot(data = blackwood_acacia_sp) +
  geom_sf(color = "darkgreen") +
  theme_minimal()

# Read in basemap shapefile
sf_map <- read_sf(here("data","sf_map","tl_2017_06075_roads.shp"))

# Match CRS
st_transform(sf_map, 4326)

# Plot
ggplot(data = sf_map) +
  geom_sf()

```

Combine data with basemap
```{r}

ggplot() +
  geom_sf(data = sf_map,
          size = 0.1,
          color = "darkgray") +
  geom_sf(data = blackwood_acacia_sp, 
          color = "red", 
          size = 0.5) +
  theme_void() +
  labs(title = "Blackwood acacias in San Francisco")

```

### Interactive! 
```{r}

tmap_mode("view")

tm_shape(blackwood_acacia_sp) + 
  tm_dots()

```
## END Lab 1
